---
title: "Assignment 03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment  

Based off of Chapter: "Data Transformation" in the R4DS book.


```{r}
library('Lahman')
library('tidyverse')
install.packages("tidyverse")
?flights
View(flights)
dat %>% 
  summarise(num_flights = n())
```


### Exercises

1.  How could you use `arrange()` to sort all missing values to the start?
    (Hint: use `is.na()`).
```{r}
dat <- flights
dat %>% 
    arrange(desc(is.na(dep_delay)),
           desc(is.na(dep_time)),
           desc(is.na(arr_delay)), 
           desc(is.na(arr_time)),
           desc(is.na(air_time)),
           desc(is.na(tailnum)))
```
    
2.  Sort `flights` to find the most delayed flights. Find the flights that
    left earliest.
```{r}
dat %>%
  arrange(desc(dep_delay)) %>%
  arrange(dep_time)
```

3.  Sort `flights` to find the fastest flights.
```{r}
dat %>%
  arrange(air_time)
```


4.  Which flights travelled the longest? Which travelled the shortest?
```{r}
dat %>%
  arrange(desc(distance)) %>% 
  arrange(distance)
```



### Exercises

1.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`,
    `arr_time`, and `arr_delay` from `flights`.
```{r}
times_and_delays <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
select(dat, times_and_delays)
dat %>% 
  select(times_and_delays)
select(dat, dep_time, dep_delay, arr_time, arr_delay)
select(dat, "dep_time", "dep_delay", "arr_time", "arr_delay")
select(dat, starts_with("dep"), starts_with("arr"))
select(dat, contains("dep"), contains("arr"), -contains("sched"), -contains("carrier"))
select(dat, ends_with("me"), ends_with("ay"), -starts_with("sched"), -starts_with("day"), -starts_with("air"))
select(dat, matches("dep"), matches("arr"), -matches("sched"), -matches("carrier"))
select(dat, matches("delay$|time$"), -contains("sched"), -contains("air"))
select(dat, matches("^dep|^arr"))

```

2.  What happens if you include the name of a variable multiple times in
    a `select()` call?
```{r}
a <- "air_time"
select(dat, rep(a, 3))
# R returns the first one but ignores the remaining.
```

3.  What does the `one_of()` function do? Why might it be helpful in conjunction
    with this vector?
    
    ```{r}
    vars <- c("year", "month", "day", "dep_delay", "arr_delay")
    #one_of() is like %in% and it checks if each colname is "one of" the given options. so it can be used as      shown below:
select(dat, one_of(vars))
    ```
    
4.  Does the result of running the following code surprise you?  How do the
    select helpers deal with case by default? How can you change that default?

    ```{r, eval = FALSE}
# Yes, it appears that R ignores case by default in these functions, so we must explicitly tell R to NOT ignore the case if we desire a specific case transformation.
    select(dat, contains("TIME", ignore.case = FALSE))
    ```




### Exercises

```{r, eval = FALSE, echo = FALSE}
View(dat)
flights <- flights %>% mutate(
  dep_time = hour * 60 + minute,
  arr_time = (arr_time %/% 100) * 60 + (arr_time %% 100),
  airtime2 = arr_time - dep_time,
  dep_sched = dep_time + dep_delay
)

ggplot(flights, aes(dep_sched)) + geom_histogram(binwidth = 60)
ggplot(flights, aes(dep_sched %% 60)) + geom_histogram(binwidth = 1)
ggplot(flights, aes(air_time - airtime2)) + geom_histogram()
```

1.  Currently `dep_time` and `sched_dep_time` are convenient to look at, but
    hard to compute with because they're not really continuous numbers. 
    Convert them to a more convenient representation of number of minutes
    since midnight.
```{r}
dat %>%
  mutate(dep_time = ((dep_time %/% 100) * 60 + (dep_time %% 100))) %>%
  mutate(sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100))
```
    
2.  Compare `air_time` with `arr_time - dep_time`. What do you expect to see?
    What do you see? What do you need to do to fix it?
```{r}
# They are different because dep_time is transformed into minutes after midnight and arr_time is in regular military clock time. The calculations won't work. We need to mutate all the times to a consistent minutes format. I get negative numbers when doing the calculations.
dat %>%
  mutate(dep_time = ((dep_time %/% 100) * 60 + (dep_time %% 100))) %>%
  mutate(arr_time = (arr_time %/% 100) * 60 + (arr_time %% 100)) %>% 
  transmute((arr_time - dep_time) %% (60*24) - air_time)
```
    
3.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you
    expect those three numbers to be related?
```{r}
# Logically, sched_dep_time + dep_time = dep_delay. We can use the same transformations as previously for dep_time and arr_time, then add a transformation for sched_arr_time and sched_dep_time to compare delays.

dat %>% 
  mutate(dep_time = (dep_time %/% 100) * 60 + (dep_time %% 100)) %>% 
  mutate(arr_time = (arr_time %/% 100) * 60 + (arr_time %% 100)) %>%         
  mutate(sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100)) %>% 
  mutate(sched_arr_time = (sched_arr_time %/% 100) * 60 + (sched_arr_time %% 100)) %>%
  transmute(near((sched_dep_time + dep_delay) %% (60*24), dep_time, tol=1))

```

4.  Find the 10 most delayed flights using a ranking function. How do you want 
    to handle ties? Carefully read the documentation for `min_rank()`.
```{r}
#?min_rank()
#?filter()
#?replace_na
dat %>% 
  filter(min_rank(desc(dep_delay))<=10) %>% 
  arrange(desc(dep_delay))
# The top 10 delayed flights do not have ties.
```

5.  What does `1:3 + 1:10` return? Why?
```{r}
1:3
1:3 + 1:10
# 1:3 creates a vector if 3 integers 1 2 3. 1:10 creates a vector of 10 integers 1 2 3 4 5 6 7 8 9 10. The addition operator "+" adds the values in each respective position in the vectors resulting in a new vector of length 10 "2  4  6  5  7  9  8 10 12 11". Because the vectors are not of equal length, the shorter vector is repeated as many times as necessary (in this case 4 times) to complete the operation. A warning is provided to warn of this. 
```

6.  What trigonometric functions does R provide?
?Trig
cos(x)
sin(x)
tan(x)

acos(x)
asin(x)
atan(x)
atan2(y, x)

cospi(x)
sinpi(x)
tanpi(x)


### Exercises

1.  Brainstorm at least 5 different ways to assess the typical delay 
    characteristics of a group of flights. Consider the following scenarios:
    
    * A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of 
      the time.
```{r}
all_delays <- flights %>%
  group_by(flight) %>%
  summarise(count_flights = n(),
            on_time_flights = sum(((arr_time == sched_arr_time)/count_flights)*100),
            early_flights_15 = sum(((sched_arr_time - arr_time == 15)/count_flights)*100),
            late_flights_15 = sum(((arr_time - sched_arr_time == 15)/count_flights)*100))

all_delays %>% 
  filter(early_flights_15 == 0.5 & late_flights_15 == 0.5)

```
      
    * A flight is always 10 minutes late.
```{r}
dat %>%
  group_by(flight) %>%
  summarise(dep_delay_mean = mean(dep_delay)) %>%
  filter(dep_delay_mean > 10) %>%
  arrange(-dep_delay_mean)
```

    * A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of 
      the time.
```{r}
all_delays <- flights %>%
  group_by(flight) %>%
  summarise(count_flights = n(),
            on_time_flights = sum(((arr_time == sched_arr_time)/count_flights)*100),
            early_flights_30 = sum(((sched_arr_time - arr_time == 30)/count_flights)*100),
            late_flights_30 = sum(((arr_time - sched_arr_time == 30)/count_flights)*100))

all_delays %>% 
  filter(early_flights_30 == 0.5 & late_flights_30 == 0.5)
```
     
    * 99% of the time a flight is on time. 1% of the time it's 2 hours late.
    
    Which is more important: arrival delay or departure delay?
```{r}
# Probably arrival delay since people have made arrangements for connecting flights, ground trnsportation, and meetings etc, requiring arrival at a specific time. 
```

2.  Come up with another approach that will give you the same output as 
    `not_cancelled %>% count(dest)` and 
    `not_cancelled %>% count(tailnum, wt = distance)` (without using 
    `count()`).
```{r}
#tally(x, wt = NULL, sort = FALSE, name = "n")
#?tally
not_cancelled <- filter(flights, !is.na(air_time)) 
cancelled <- filter(flights, is.na(air_time))
not_cancelled %>% group_by(dest) %>% tally()
not_cancelled %>% group_by(dest) %>% tally(wt=distance)
```

3.  Our definition of cancelled flights (`is.na(dep_delay) | is.na(arr_delay)`
    ) is slightly suboptimal. Why? Which is the most important column?
```{r}
#Because dep_delay and arr_delay could be recorded even if a flight was cancelled; whereas, air_time only is recorded when a flight has actually been airbourne. air_time is the best column to determine cancelled / non_cancelled flights.
```

4.  Look at the number of cancelled flights per day. Is there a pattern?
    Is the proportion of cancelled flights related to the average delay?
```{r}
#dat %>% 
#  select(day)
#cancelled %>% group_by(day) %>% count()
dat %>% group_by(day) %>%
  summarise(cancelled = mean(is.na(dep_delay)),
            mean_dep = mean(dep_delay, na.rm = T),
            mean_arr = mean(arr_delay, na.rm = T)) %>%
  ggplot(aes(y = cancelled)) +
  geom_point(aes(x = mean_dep), colour = "red") +
  geom_point(aes(x = mean_arr), colour = "blue") +
  labs(x = "Avg delay per day", y = "Cancelled flights p day")
```

5.  Which carrier has the worst delays? Challenge: can you disentangle the
    effects of bad airports vs. bad carriers? Why/why not? (Hint: think about
    `flights %>% group_by(carrier, dest) %>% summarise(n())`)

6.  What does the `sort` argument to `count()` do. When might you use it?



### Exercises

1.  Refer back to the lists of useful mutate and filtering functions. 
    Describe how each operation changes when you combine it with grouping.

2.  Which plane (`tailnum`) has the worst on-time record?

3.  What time of day should you fly if you want to avoid delays as much
    as possible?
    
4.  For each destination, compute the total minutes of delay. For each 
    flight, compute the proportion of the total delay for its destination.
    
5.  Delays are typically temporally correlated: even once the problem that
    caused the initial delay has been resolved, later flights are delayed 
    to allow earlier flights to leave. Using `lag()`, explore how the delay
    of a flight is related to the delay of the immediately preceding flight.
    
6.  Look at each destination. Can you find flights that are suspiciously
    fast? (i.e. flights that represent a potential data entry error). Compute
    the air time a flight relative to the shortest flight to that destination.
    Which flights were most delayed in the air?
    
7.  Find all destinations that are flown by at least two carriers. Use that
    information to rank the carriers.

8.  For each plane, count the number of flights before the first delay 
    of greater than 1 hour.
